ASAH1 rGBM Manuscript ELISA Figure

### 2023-01-03 results/elisa_plasma_proteomics/manuscript_elisa_figure_final.Rmd
[ADD] Update figures and redo ratio sPLS-DA analysis

## Build the plot ----
colfunc <- colorRampPalette(c("red","yellow","springgreen","royalblue"))
ttr_categorizer <- function(data) {
  data <- as.numeric(data)
  if (is.na(data)){
    return(NA)
  }
  if (data >= 100 & data < 200){
    return("100-200")
  } else if (data >= 200 & data < 300){
    return("200-300")
  } else if (data >= 300 & data < 400){
    return("300-400")
  } else if (data >= 400 & data < 500){
    return("400-500")
  } else if (data >= 500 & data < 600){
    return("500-600")
  }
}
# ttr_categorizer <- function(data) {
#   data <- as.numeric(data)
#   if (is.na(data)){
#     return(NA)
#   }
#   if (data >= 100 & data < 150){
#     return("100-150")
#   } else if (data >= 150 & data < 200){
#     return("150-200")
#   } else if (data >= 200 & data < 250){
#     return("200-250")
#   } else if (data >= 250 & data < 300){
#     return("250-300")
#   } else if (data >= 300 & data < 350){
#     return("300-350")
#   } else if (data >= 350 & data < 400){
#     return("350-400")
#   } else if (data >= 400 & data < 450){
#     return("400-450")
#   } else if (data >= 450 & data < 500){
#     return("450-500")
#   } else if (data >= 500 & data < 550){
#     return("500-550")
#   } else if (data >= 550 & data < 600){
#     return("550-600")
#   }
# }
# ttr_categorizer <- function(data, split_value) {
#   if (data <= split_value) {
#     return(0)
#   } else {
#     return(1)
#   }
# }
# ttr_categorizer <- function(data) {
#   if (data >= 103 & data < 190){
#     return("early")
#   } else if (data >= 190 & data < 243){
#     return("middle")
#   } else if (data >= 243 & data <= 521){
#     return("late")
#   }
# }
pdata$`ttr_group` <- sapply(as.numeric(pdata$`Time-to-reccurence (days)`), ttr_categorizer)
ggpaired(data = pdata, id = "Number.x", x = "Manifestation", y = "log2_conc", fill = "Manifestation", palette = colfunc(length(unique(pdata$`ttr_group`))), line.color = "ttr_group", line.size = 0.4) +
  # stat_compare_means(paired = TRUE, label.x = 1.5, geom = "label") +
  # stat_compare_means(comparisons = my_comparisons, group.by="gene", label.y = c(4, 7, 7), label.x = c(1, 2, 3)) +
  scale_fill_manual(values = RColorBrewer::brewer.pal(length(unique(pdata$Manifestation)), "Dark2"), labels = c("Ctrl", "Initial manifestation", "Second manifestation")) +
  # geom_jitter(aes(label=Number.x)) +
  # ggtitle("Initial vs Recurrence") +
  labs(x="", y="log2( Concentration )") +
  facet_wrap(~gene, scales = "free_y") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size=17),
        axis.title = element_text(size = 17),
        axis.text = element_text(size=16),
        strip.text = element_text(size=16),
        legend.title = element_text(size=16),
        legend.text = element_text(size=15),
        legend.position = "bottom", legend.box="vertical",) +
  guides(fill=guide_legend(title="GBM\nOccurence", ncol=1),
         col=guide_legend(title="Time to Recurrence")) -> p1

p1 + 
  stat_pvalue_manual(stat_test, label = "{p.signif} p = {p}", step.group.by = "gene", step.increase = 0.2, hide.ns = T, size=4, bracket.size=1) +
  geom_point(data=data.frame(Manifestation=c("iGBM", "iGBM", "iGBM", "iGBM"), gene=c("ASAH1", "SYNEMIN", "GPNMB", "MMP-9" ), y=c(8, 18, 9, 12)),
             aes(x=Manifestation, y=y, group=gene), alpha=0) -> p1
# p1

@@

## Build the plot ----
ggpaired(data=pdata2, id = "Number.x", x = "collection_type2", y = "log2_conc", fill = "collection_type2", palette = "jco", line.color = "grey", line.size = 0.4) +
  scale_fill_manual(values = RColorBrewer::brewer.pal(length(unique(pdata2$`collection_type2`)), "Dark2"), labels = c("Ctrl", "pre-surgery", "post-surgery")) +
  # stat_compare_means(paired = TRUE) +
  # ggnewscale::new_scale_color() +
  # geom_hline(data=mean_ctrl, aes(yintercept=mean_conc, col="ctrl"), linetype="dashed") +
  # geom_rect(data=mean_ctrl, aes(xmin=-Inf, xmax=Inf, ymin=mean_conc-sd_conc, ymax=mean_conc+sd_conc), fill="red", alpha=0.1) +
  # scale_color_manual(label="Mean Conc.", values="red") +
  # ggtitle("Pre-Surgery vs Post-Surgery") +
  labs(x="", y="log2( Concentration )") +
  facet_wrap(~gene, scales = "free_y") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size=17),
        axis.title = element_text(size = 17),
        axis.text = element_text(size=16),
        strip.text = element_text(size=16),
        legend.title = element_text(size=16),
        legend.text = element_text(size=15),
        legend.position = "bottom", legend.box="vertical",) +
  guides(fill=guide_legend(title="EDTA Blood\nCollection Type", ncol=1),
         col=guide_legend(title="Healthy Control Baseline")) -> p2

@@

@@

@@

master_roc_data <- rbindlist(list(`Ctrl vs iGBM`=pred_obj_1.1$roc_data, `Ctrl vs rGBM`=pred_obj_1.2$roc_data, `iGBM vs rGBM protein`=tmp_pred_obj_1.5_roc_data, `iGBM vs rGBM protein + clinical`=tmp_pred_obj_1.4_roc_data, `iGBM vs rGBM protein ratios`=tmp_pred_obj_1.3_roc_data), idcol = "Comparison")

@@

@@

g <- ggplot(master_roc_data[Model %in% c("Elastic Logit")], aes(m=GBM, d=factor(obs, levels = c("ctrl", "GBM")), color=Comparison, linetype=Comparison)) +
  geom_roc(n.cuts=0, size=1.5) + 
  scale_color_manual(values = RColorBrewer::brewer.pal(5, "Dark2")) +
  # scale_linetype_manual(values=c("solid", "dotted"), labels=c("Elastic Logit")) + 
  coord_equal() +
  style_roc(x="1 - Specificity", y="Sensitivity") +
  annotation_custom(auc_grob, xmin=0.23, xmax=0.90, ymin=0.10, ymax=0.38) +
  theme(axis.title = element_text(size=14),
        axis.text = element_text(size=13),
        legend.title = element_text(size=14),
        legend.text = element_text(size=13))

ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_d.png", plot = g, width = 11, height = 8.5, dpi = 200)
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_d.svg", plot = g, width = 11, height = 8.5, dpi = 200)
subfigure_1 <- g
```

@@

fit <- survfit(Surv(Latency..days., status) ~ ASAH1, data = latency_data)
print(fit)

# Change color, linetype by strata, risk.table color by strata
surv_p1 <- ggsurvplot(fit,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          ) +
  labs(x="Latency Time (Days)")

fit <- survfit(Surv(Latency..days., status) ~ SYNM, data = latency_data)
print(fit)

# Change color, linetype by strata, risk.table color by strata
surv_p2 <- ggsurvplot(fit,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          ) +
  labs(x="Latency Time (Days)")

fit <- survfit(Surv(Latency..days., status) ~ GPNMB, data = latency_data)
print(fit)

# Change color, linetype by strata, risk.table color by strata
surv_p3 <- ggsurvplot(fit,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          ) +
  labs(x="Latency Time (Days)")

>>>>

## Build the plot ----
colfunc <- colorRampPalette(c("red","yellow","springgreen","royalblue"))
ttr_categorizer <- function(data) {
  data <- as.numeric(data)
  if (is.na(data)){
    return(NA)
  }
  if (data >= 100 & data < 200){
    return("100-200")
  } else if (data >= 200 & data < 300){
    return("200-300")
  } else if (data >= 300 & data < 400){
    return("300-400")
  } else if (data >= 400 & data < 500){
    return("400-500")
  } else if (data >= 500 & data < 600){
    return("500-600")
  }
}
# ttr_categorizer <- function(data) {
#   data <- as.numeric(data)
#   if (is.na(data)){
#     return(NA)
#   }
#   if (data >= 100 & data < 150){
#     return("100-150")
#   } else if (data >= 150 & data < 200){
#     return("150-200")
#   } else if (data >= 200 & data < 250){
#     return("200-250")
#   } else if (data >= 250 & data < 300){
#     return("250-300")
#   } else if (data >= 300 & data < 350){
#     return("300-350")
#   } else if (data >= 350 & data < 400){
#     return("350-400")
#   } else if (data >= 400 & data < 450){
#     return("400-450")
#   } else if (data >= 450 & data < 500){
#     return("450-500")
#   } else if (data >= 500 & data < 550){
#     return("500-550")
#   } else if (data >= 550 & data < 600){
#     return("550-600")
#   }
# }
# ttr_categorizer <- function(data, split_value) {
#   if (data <= split_value) {
#     return(0)
#   } else {
#     return(1)
#   }
# }
# ttr_categorizer <- function(data) {
#   if (data >= 103 & data < 190){
#     return("early")
#   } else if (data >= 190 & data < 243){
#     return("middle")
#   } else if (data >= 243 & data <= 521){
#     return("late")
#   }
# }
pdata$`ttr_group` <- sapply(as.numeric(pdata$`Time-to-reccurence (days)`), ttr_categorizer)
ggpaired(data = pdata, id = "Number.x", x = "Manifestation", y = "log2_conc", fill = "Manifestation", palette = 'grey', line.color = "grey", line.size = 0.4) +
  # stat_compare_means(paired = TRUE, label.x = 1.5, geom = "label") +
  # stat_compare_means(comparisons = my_comparisons, group.by="gene", label.y = c(4, 7, 7), label.x = c(1, 2, 3)) +
  scale_fill_manual(values = RColorBrewer::brewer.pal(length(unique(pdata$Manifestation)), "Dark2"), labels = c("Ctrl", "iGBM", "rGBM")) +
  # geom_jitter(aes(label=Number.x)) +
  # ggtitle("Initial vs Recurrence") +
  labs(x="", y="log2( Concentration )") +
  facet_wrap(~gene, scales = "free_y") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size=17),
        axis.title = element_text(size = 17),
        axis.text = element_text(size=16),
        strip.text = element_text(size=16),
        legend.title = element_text(size=16),
        legend.text = element_text(size=15),
        legend.position = "bottom", legend.box="vertical",) +
  guides(fill=guide_legend(title="", ncol=3),
         col='none') -> p1_org

p1_org + 
  stat_pvalue_manual(stat_test, label = "{p.signif} p = {p}", step.group.by = "gene", step.increase = 0.2, hide.ns = T, size=4, bracket.size=1) +
  geom_point(data=data.frame(Manifestation=c("iGBM", "iGBM", "iGBM", "iGBM"), gene=c("ASAH1", "SYNEMIN", "GPNMB", "MMP-9" ), y=c(8, 18, 9, 12)),
             aes(x=Manifestation, y=y, group=gene), alpha=0) -> p1_org

# Exclude outlier patient number.x 54 
ggpaired(data = pdata[Number.x!=54], id = "Number.x", x = "Manifestation", y = "log2_conc", fill = "Manifestation", palette = 'grey', line.color = "grey", line.size = 0.4) +
  # stat_compare_means(paired = TRUE, label.x = 1.5, geom = "label") +
  # stat_compare_means(comparisons = my_comparisons, group.by="gene", label.y = c(4, 7, 7), label.x = c(1, 2, 3)) +
  scale_fill_manual(values = RColorBrewer::brewer.pal(length(unique(pdata$Manifestation)), "Dark2"), labels = c("Ctrl", "iGBM", "rGBM")) +
  # geom_jitter(aes(label=Number.x)) +
  # ggtitle("Initial vs Recurrence") +
  labs(x="", y="log2( Concentration )") +
  facet_wrap(~gene, scales = "free_y") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size=17),
        axis.title = element_text(size = 17),
        axis.text = element_text(size=16),
        strip.text = element_text(size=16),
        legend.title = element_text(size=16),
        legend.text = element_text(size=15),
        legend.position = "bottom", legend.box="vertical",) +
  guides(fill=guide_legend(title="", ncol=3),
         col='none') -> p1

p1 + 
  stat_pvalue_manual(stat_test, label = "{p.signif} p = {p}", step.group.by = "gene", step.increase = 0.2, hide.ns = T, size=4, bracket.size=1) +
  geom_point(data=data.frame(Manifestation=c("iGBM", "iGBM", "iGBM", "iGBM"), gene=c("ASAH1", "SYNEMIN", "GPNMB", "MMP-9" ), y=c(7, 15, 8, 11)),
             aes(x=Manifestation, y=y, group=gene), alpha=0) -> p1
# p1

@@

## Build the plot ----
ggpaired(data=pdata2, id = "Number.x", x = "collection_type2", y = "log2_conc", fill = "collection_type2", palette = "jco", line.color = "grey", line.size = 0.4) +
  scale_fill_manual(values = RColorBrewer::brewer.pal(length(unique(pdata2$`collection_type2`)), "Dark2"), labels = c("Ctrl", "pre-surgery", "post-surgery")) +
  # stat_compare_means(paired = TRUE) +
  # ggnewscale::new_scale_color() +
  # geom_hline(data=mean_ctrl, aes(yintercept=mean_conc, col="ctrl"), linetype="dashed") +
  # geom_rect(data=mean_ctrl, aes(xmin=-Inf, xmax=Inf, ymin=mean_conc-sd_conc, ymax=mean_conc+sd_conc), fill="red", alpha=0.1) +
  # scale_color_manual(label="Mean Conc.", values="red") +
  # ggtitle("Pre-Surgery vs Post-Surgery") +
  labs(x="", y="log2( Concentration )") +
  facet_wrap(~gene, scales = "free_y") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size=17),
        axis.title = element_text(size = 17),
        axis.text = element_text(size=16),
        strip.text = element_text(size=16),
        legend.title = element_text(size=16),
        legend.text = element_text(size=15),
        legend.position = "bottom", legend.box="vertical",) +
  guides(fill=guide_legend(title="EDTA Blood\nCollection Type", ncol=3),
         col=guide_legend(title="Healthy Control Baseline")) -> p2

@@

subfigure_6_org <- p1_org

@@

# Subsetting and hold out patient for Validating Ratios

I still think there is value in usiing the ratios, so I performed the sPLS-DA on the ratio data again, but instead of using all the ratios and inverse ratios from all patients, I do the following:

1. Randomly sample a single patient to use as a test patient, to make predictions on
2. Randomly sample n random patients from remaining patients and compute iGBM/rGBM ratios
3. For the rest of the patients I compute rGBM/iGBM ratios

This results in the following data used in the sPLS-DA model

**Training Data**

| Patient ID | Class | ASAH1      | SYNMEIN    | GPNMB      | MMP-9      |
|------------|-------|------------|------------|------------|------------|
| 6          | iGBM  | -4.3219281 | 2.6629410  | 0.3162515  | 3.8886097  |
| 10         | rGBM  | 2.2103599  | -3.8779037 | 2.5628539  | -3.1538943 |
| 16         | rGBM  | 1.6956066  | 0.8506559  | 0.1659252  | -1.9814540 |
| 23         | rGBM  | -0.0257631 | -3.6121574 | 1.6151451  | -1.7166787 |
| 26         | iGBM  | 1.7744403  | 0.4268049  | 4.2786146  | -1.5475206 |
| 29         | rGBM  | 3.3645724  | 0.4959143  | -0.6652013 | 0.5074684  |
| 31         | iGBM  | -1.6599246 | -2.8570919 | -1.9497226 | -1.3115642 |
| 34         | iGBM  | -0.1768778 | 0.7865737  | 2.1875247  | 1.1604031  |

**Test Data**
| Patient ID | Class | ASAH1      | SYNMEIN    | GPNMB      | MMP-9      |
|------------|-------|------------|------------|------------|------------|
| 35          | rGBM  | -0.02599521 | -4.115453  | -2.892362  | -1.679008  |

This way we do not duplicate patient IDs (i.e. having iGBM/rGBM and rGBM/iGBM for the same patient) and we have a single patient we can make a prediction on to see where in the PLS-DA dimension they would land.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=11, fig.width=21, include=FALSE}
data_long2 <- data_long

## Data wrangling ----
### iGBM over rGBM ratio
data_long2 %>% 
  dplyr::mutate(conc = conc+1) %>% # Add one to avoid cases of 0 division
  dplyr::group_by(gene, Number.x) %>%
  dplyr::group_map(function(.d, ...){
    .d <- as.data.table(.d)
    .d %>% 
      dplyr::group_by(gene, Number.x, patient_state, `Age at surgery (years)`, Sex, `Tumor localization`, `Type of resection`, `MGMT promotor methylation`, `p53 accumulation`, `Ki67-Li`, `EGFR vIII`, Treatment, `Time-to-reccurence (days)`, `Latency (days)`) %>% 
      dplyr::summarise(agg_lab_id = paste(LAB_ID, collapse=";"),
                       agg_collection_type = paste(collection_type, collapse = ";"), .groups = "keep") -> tmp
    # print(.d)
    if ( nrow(.d)==2 & any(!is.na(.d$manifestation_type)) ){
      .d_initial <- .d[manifestation_type=="Initial manifestation"]
      .d_second <- .d[manifestation_type=="Second manifestation"]
      tmp$ratio <- .d_initial$conc / .d_second$conc
    } else if (nrow(.d)==2) {
      .d_initial <- .d[collection_type=="E1"]
      .d_second <- .d[collection_type=="E2"]
      tmp$ratio <- .d_initial$conc / .d_second$conc
    } else {
      tmp$ratio <- NA #.d$conc
    }
    return(tmp)
  }, .keep = TRUE) %>%
  rbindlist() %>%
  dplyr::mutate(sample_comparison = plyr::mapvalues(patient_state, from=c(0, 1, 2), to=c("ctrl", "iGBMoverrGBM", "pre-surgery/post-surgery"))) %>%
  dplyr::filter(sample_comparison=="iGBMoverrGBM" ) %>%
  dplyr::mutate(log2_conc = log2(ratio)) %>%
  # dplyr::group_by(gene, Number.x) %>%
  # dplyr::add_count() %>%
  # dplyr::filter(n==2) %>%
  dplyr::mutate(class = plyr::mapvalues(patient_state, from=c(1), to=c("iGBMoverrGBM"))) -> pdata

### rGBM over iGBM
data_long2 %>% 
  dplyr::mutate(conc = conc+1) %>% # Add one to avoid cases of 0 division
  dplyr::filter(!is.na(manifestation_type) | patient_state!=0) %>%
  dplyr::group_by(gene, Number.x) %>%
  dplyr::group_map(function(.d, ...){
    .d <- as.data.table(.d)
    .d %>% 
      dplyr::group_by(gene, Number.x, patient_state, `Age at surgery (years)`, Sex, `Tumor localization`, `Type of resection`, `MGMT promotor methylation`, `p53 accumulation`, `Ki67-Li`, `EGFR vIII`, Treatment, `Time-to-reccurence (days)`, `Latency (days)`) %>% 
      dplyr::summarise(agg_lab_id = paste(LAB_ID, collapse=";"),
                       agg_collection_type = paste(collection_type, collapse = ";"), .groups = "keep") -> tmp
    # print(.d)
    if ( nrow(.d)==2 & any(!is.na(.d$manifestation_type)) ){
      .d_initial <- .d[manifestation_type=="Initial manifestation"]
      .d_second <- .d[manifestation_type=="Second manifestation"]
      tmp$ratio <- .d_second$conc / .d_initial$conc
    } else if (nrow(.d)==2) {
      .d_initial <- .d[collection_type=="E1"]
      .d_second <- .d[collection_type=="E2"]
      tmp$ratio <- .d_initial$conc / .d_second$conc
    } else {
      tmp$ratio <- NA #.d$conc
    }
    return(tmp)
  }, .keep = TRUE) %>%
  rbindlist() %>%
  dplyr::mutate(sample_comparison = plyr::mapvalues(patient_state, from=c(0, 1, 2), to=c("ctrl", "rGBMoveriGBM", "pre-surgery/post-surgery"))) %>%
  dplyr::filter(sample_comparison=="rGBMoveriGBM" ) %>%
  dplyr::filter(grepl(";", agg_lab_id)) %>%
  dplyr::mutate(log2_conc = log2(ratio)) %>%
  # dplyr::group_by(gene, Number.x) %>%
  # dplyr::add_count() %>%
  # dplyr::filter(n==2) %>%
  dplyr::mutate(class = plyr::mapvalues(patient_state, from=c(1), to=c("rGBMoveriGBM"))) -> pdata2

## Remove non-paired patients
data_wide %>%
  dplyr::group_by(Number.x) %>%
  dplyr::add_count() %>%
  dplyr::filter(n==2) %>%
  dplyr::ungroup() %>%
  dplyr::select(-n) -> data_wide

# Sample for Test data
test_patients <- sample(unique(train_df$Number.x), 1)
test_df <- train_df %>%
  dplyr::filter(Number.x %in% test_patients) %>%
  dplyr::filter(class=="rGBM")

# Refine Training data
train_df %>%
  dplyr::filter(!(Number.x %in% test_patients)) -> train_df
# Sample patients to select half patients for iGMB/rGBM and half patients for rGBM/iGBM
igbm_train_patients <- sample(unique(train_df$Number.x), length(unique(train_df$Number.x))/2)
rgbm_train_patients <- unique(train_df$Number.x)[ !unique(train_df$Number.x) %in% igbm_train_patients]

train_df %>%
  dplyr::filter((Number.x %in% igbm_train_patients & class=='iGBM') | (Number.x %in% rgbm_train_patients & class=='rGBM')) -> train_df

MyResult.splsda <- mixOmics::splsda(train_df[, c(-1, -2, -7, -8, -9, -10)], train_df$class, keepX = c(4,4))

pred_test <- predict(MyResult.splsda, test_df[, c(-1, -2, -7, -8, -9, -10)], test_df$class)

background <- background.predict(MyResult.splsda, comp.predicted=2, dist = "max.dist", xlim = c(min(c(MyResult.splsda$variates$X, pred_test$variates)), max(c(MyResult.splsda$variates$X, pred_test$variates))), ylim = c(min(c(MyResult.splsda$variates$X, pred_test$variates)), max(c(MyResult.splsda$variates$X, pred_test$variates))))

plsda_data <- cbind( patient_id=na.omit(data_wide %>% dplyr::filter(!(Number.x %in% test_patients)))  %>% dplyr::pull(Number.x), class=train_df$class, as.data.frame(MyResult.splsda$variates$X))
plsda_data$patient_id <- as.factor(plsda_data$patient_id)
plsda_data %>%
  dplyr::arrange(class) -> plsda_data
plsda_data <- as.data.table(plsda_data)

p <- ggplot() +
  geom_polygon(data = background,aes(x=Var1, y=Var2,
                                     fill = col), inherit.aes = FALSE, show.legend=FALSE, alpha=0.2) +
  scale_fill_manual(values=c("#388ECC", "#F68B33")) +
  # geom_point(data=plsda_data[patient_id=="Ctrl"], aes(x=comp1, y=comp2, shape=class), col="black", size=3) +
  # geom_point(data=plsda_data[patient_id!="Ctrl"], aes(x=comp1, y=comp2, col=class, shape=class), size=3) +
  geom_point(data=plsda_data, aes(x=comp1, y=comp2, col=class, shape=class), size=3) +
  scale_shape_manual(values=c(15, 16, 17)) +
  scale_color_manual(values=c( "#388ECC", "#F68B33"), labels=c(expression(paste(frac('initial GBM', 'recurrent GBM'), " Ratio")),
                                                               expression(paste(frac('recurrent GBM', 'initial GBM'), " Ratio")))
  ) +
  ggnewscale::new_scale_color() +
  geom_path(data=average_comp, aes(x=avg_comp1, y=avg_comp2, group = patient_id, col="Avg. iGBM\nto rGBM\nTrajectory"),
            arrow = arrow(type = "closed", ends='both',
                          length=unit(0.1, "inches")), size=2, alpha=0.8) +
  ggnewscale::new_scale_color() +
  geom_point(data=data.frame(x=pred_test$variates[,1], y=pred_test$variates[,2]), aes(x=x, y=y, col="black"), size=3, shape=3) +
  scale_color_manual(values=c( "black"), labels=c(expression(paste(frac('recurrent GBM', 'initial GBM'), " Ratio")))) +
  labs(x=sprintf("X-variate 1: %s%% explained variance", round(MyResult.splsda$prop_expl_var$X[1]*100)), y =sprintf("X-variate 2: %s%% explained variance", round(MyResult.splsda$prop_expl_var$X[2]*100))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size=17),
        axis.title = element_text(size = 17),
        axis.text = element_text(size=16),
        strip.text = element_text(size=16),
        legend.title = element_text(size=16),
        legend.text = element_text(size=15),
        legend.position = "right", legend.box="vertical",) +
  guides(shape="none",
         colour_new_new=guide_legend(title="Class", ncol=1, byrow = T),
         colour_new=guide_legend(title="Trajectory", ncol=1, byrow = T),
         col=guide_legend(title="Prediction on\nHeld Out Patients", ncol=1))

splsda_better_ratio_avg_traj_fig <- p
splsda_better_ratio_avg_traj_fig
# ggsave("../../figures/elisa_plasma_proteomics/elisa_supp_figure3_subfigure_b.png", plot = splsda_ratio_avg_traj_fig, width = 11, height = 8.5, dpi = 200)
# ggsave("../../figures/elisa_plasma_proteomics/elisa_supp_figure3_subfigure_b.svg", plot = splsda_ratio_avg_traj_fig, width = 11, height = 8.5, dpi = 200)
# ggsave("../../figures/elisa_plasma_proteomics/elisa_supp_subfigure_b.png", plot = splsda_ratio_avg_traj_fig, width = 11, height = 8.5, dpi = 200)
# ggsave("../../figures/elisa_plasma_proteomics/elisa_supp_subfigure_b.svg", plot = splsda_ratio_avg_traj_fig, width = 11, height = 8.5, dpi = 200)

# p <- ggplot() +
#   geom_polygon(data = background,aes(x=Var1, y=Var2,
#                                      fill = col), inherit.aes = FALSE, show.legend=FALSE, alpha=0.2) +
#   scale_fill_manual(values=c( "#388ECC", "#F68B33")) +
#   # geom_point(data=plsda_data[patient_id=="Ctrl"], aes(x=comp1, y=comp2, shape=class), col="black", size=3) +
#   # geom_point(data=plsda_data[patient_id!="Ctrl"], aes(x=comp1, y=comp2, col=class, shape=class), size=3) +
#   geom_point(data=plsda_data, aes(x=comp1, y=comp2, col=class, shape=class), size=3) +
#   scale_shape_manual(values=c(15, 16, 17)) +
#   scale_color_manual(values=c( "#388ECC", "#F68B33"), labels=c(expression(paste(frac('initial GBM', 'recurrent GBM'), " Ratio")),
#                                                                expression(paste(frac('recurrent GBM', 'initial GBM'), " Ratio")))
#   ) +
#   ggnewscale::new_scale_color() +
#   geom_path(data=plsda_data[patient_id!="Ctrl"], aes(x=comp1, y=comp2, group = patient_id, col=patient_id),
#             arrow = arrow(type = "closed", ends='both',
#                           length=unit(0.1, "inches")), size=1.5, alpha=0.8) +
#   ggnewscale::new_scale_color() +
#   geom_point(data=data.frame(x=pred_test$variates[,1], y=pred_test$variates[,2]), aes(x=x, y=y, col="black"), size=3, shape=3) +
#   scale_color_manual(values=c( "black"), labels=c(expression(paste(frac('recurrent GBM', 'initial GBM'), " Ratio")))) +
#   labs(x=sprintf("X-variate 1: %s%% explained variance", round(MyResult.splsda$prop_expl_var$X[1]*100)), y =sprintf("X-variate 2: %s%% explained variance", round(MyResult.splsda$prop_expl_var$X[2]*100))) +
#   theme_classic() +
#   theme(plot.title = element_text(hjust = 0.5, size=17),
#         axis.title = element_text(size = 17),
#         axis.text = element_text(size=16),
#         strip.text = element_text(size=16),
#         legend.title = element_text(size=16),
#         legend.text = element_text(size=15),
#         legend.position = "right", legend.box="vertical",) +
#   guides(shape="none",
#          colour_new_new=guide_legend(title="Class", ncol=1, byrow = T),
#          colour_new=guide_legend(title="Patient ID\nTrajectory", ncol=2),         
#          col=guide_legend(title="Prediction on\nHeld Out Patients", ncol=1))
# 
# splsda_randomized_ratio_ind_traj_fig <- p
# splsda_randomized_ratio_ind_traj_fig
# ggsave("../../figures/elisa_plasma_proteomics/elisa_supp_figure3_subfigure_c.png", plot = splsda_ratio_ind_traj_fig, width = 11, height = 8.5, dpi = 200)
# ggsave("../../figures/elisa_plasma_proteomics/elisa_supp_figure3_subfigure_c.svg", plot = splsda_ratio_ind_traj_fig, width = 11, height = 8.5, dpi = 200)

@@

# Full
master_roc_data <- rbindlist(list(`Ctrl vs iGBM`=pred_obj_1.1$roc_data, `Ctrl vs rGBM`=pred_obj_1.2$roc_data, `iGBM vs rGBM protein`=tmp_pred_obj_1.5_roc_data, `iGBM vs rGBM protein + clinical`=tmp_pred_obj_1.4_roc_data, `iGBM vs rGBM protein ratios`=tmp_pred_obj_1.3_roc_data), idcol = "Comparison")

@@

# Only iGMB and rGBM
master_roc_data <- rbindlist(list(`iGBM vs rGBM protein`=tmp_pred_obj_1.5_roc_data, `iGBM vs rGBM protein + clinical`=tmp_pred_obj_1.4_roc_data, `iGBM vs rGBM protein ratios`=tmp_pred_obj_1.3_roc_data), idcol = "Comparison")

master_auc_data <- rbindlist(list(`iGBM vs rGBM protein`=auc_tbl_1.5, `iGBM vs rGBM protein + clinical`=auc_tbl_1.4, `iGBM vs rGBM protein ratios`=auc_tbl_1.3), idcol = "Comparison")

@@

g <- ggplot(master_roc_data[Model %in% c("Elastic Logit")], aes(m=GBM, d=factor(obs, levels = c("ctrl", "GBM")), color=Comparison), linetype='solid') +
  geom_roc(n.cuts=0, size=1.5) + 
  scale_color_manual(values = RColorBrewer::brewer.pal(5, "Dark2")) +
  # scale_linetype_manual(values=c("solid", "dotted"), labels=c("Elastic Logit")) + 
  coord_equal() +
  style_roc(x="1 - Specificity", y="Sensitivity") +
  annotation_custom(auc_grob, xmin=0.25, xmax=0.95, ymin=0.10, ymax=0.30) +
  theme(axis.title = element_text(size=14),
        axis.text = element_text(size=13),
        legend.title = element_text(size=14),
        legend.text = element_text(size=13))

ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_d.png", plot = g, width = 11, height = 8.5, dpi = 200)
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_d.svg", plot = g, width = 11, height = 8.5, dpi = 200)
subfigure_1 <- g

## Separate ROCs for controls

master_roc_data <- rbindlist(list(`Ctrl vs iGBM`=pred_obj_1.1$roc_data, `Ctrl vs rGBM`=pred_obj_1.2$roc_data), idcol = "Comparison")

master_auc_data <- rbindlist(list(`Ctrl vs iGBM`=auc_tbl_1.1, `Ctrl vs rGBM`=auc_tbl_1.2), idcol = "Comparison")

g <- ggplot(master_roc_data[Model %in% c("Elastic Logit")], aes(m=GBM, d=factor(obs, levels = c("ctrl", "GBM")), color=Comparison), linetype='solid') +
  geom_roc(n.cuts=0, size=1.5) + 
  scale_color_manual(values = RColorBrewer::brewer.pal(5, "Dark2")) +
  # scale_linetype_manual(values=c("solid", "dotted"), labels=c("Elastic Logit")) + 
  coord_equal() +
  style_roc(x="1 - Specificity", y="Sensitivity") +
  annotation_custom(auc_grob, xmin=0.23, xmax=0.90, ymin=0.10, ymax=0.38) +
  theme(axis.title = element_text(size=14),
        axis.text = element_text(size=13),
        legend.title = element_text(size=14),
        legend.text = element_text(size=13))

# ggsave("../../figures/elisa_plasma_proteomics/elisa_roc_controls.png", plot = g, width = 11, height = 8.5, dpi = 200)
# ggsave("../../figures/elisa_plasma_proteomics/elisa_roc_controls.svg", plot = g, width = 11, height = 8.5, dpi = 200)
roc_controls <- g
```

@@

fit <- survfit(Surv(Latency..days., status) ~ ASAH1, data = latency_data)
print(fit)
n_data <- table(ttr_surv_data[!is.na(ttr_surv_data$Latency..days.)]$ASAH1)
# Change color, linetype by strata, risk.table color by strata
surv_p1 <- ggsurvplot(fit,
          pval = TRUE, pval.coord = c(1700, 0.95), conf.int = TRUE,
          risk.table = FALSE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2",
          legend.title = "",
          legend.labs = c(sprintf("ASAH1=high (n=%s)", n_data[1]), sprintf("ASAH1=low (n=%s)", n_data[2]))
          ) +
  labs(x="Latency Time (Days)")

surv_p1 <- surv_p1$plot +
  theme(axis.title = element_text(size=16),
        axis.text = element_text(size=15),
        legend.text = element_text(size=13))
